---
title: "cancer cancer survival analysis (female vs male)"
output:
  pdf_document: default
  html_notebook: default
---

## Format
inst:	Institution code
time:	Survival time in days

status:	censoring status 1=censored, 2=dead

age:	Age in years

sex:	Male=1 Female=2

ph.ecog:	ECOG performance score as rated by the physician. 0=asymptomatic, 1= symptomatic but completely ambulatory, 2= in bed <50% of the day, 3= in bed > 50% of the day but not bedbound, 4 = bedbound

ph.karno:	Karnofsky performance score (bad=0-good=100) rated by physician

pat.karno:	Karnofsky performance score as rated by patient

meal.cal:	Calories consumed at meals

wt.loss:	Weight loss in last six months (pounds)

## Note
The use of 1/2 for alive/dead instead of the usual 0/1 is a historical footnote. For data contained on punch cards, IBM 360 Fortran treated blank as a zero, which led to a policy within the section of Biostatistics to never use "0" as a data value since one could not distinguish it from a missing value. The policy became a habit, as is often the case; and the 1/2 coding endured long beyond the demise of punch cards and Fortran.

## Source
Terry Therneau

## References
Loprinzi CL. Laurie JA. Wieand HS. Krook JE. Novotny PJ. Kugler JW. Bartel J. Law M. Bateman M. Klatt NE. et al. Prospective evaluation of prognostic variables from patient-completed questionnaires. North Central Cancer Treatment Group. Journal of Clinical Oncology. 12(3):601-7, 1994.

```{r}
library(dplyr)
library(DataExplorer)
library(survival)
library(survminer)
library(gmodels)
library(ggplot2)
library(tidyverse)
library(readxl)
library(epiDisplay)
library(pivottabler)
library(arsenal)
```


```{r}
# Load the survival package
library(survival)

# Load the cancer dataset from the survival package
data(cancer, package = "survival")
```


## Exploratory Data Analysis

```{r}
# Dimensions of the dataset
dim(cancer)

# Structure of the dataset
str(cancer)

# First few rows of the dataset
head(cancer)

# Concise summary of the dataset
glimpse(cancer)

# Summary statistics
summary(cancer)

```
Inst, status, sex and ph.ecog --> factor

```{r}
#Change numeric variable to factor
cancer$inst <- as.factor(cancer$inst)
cancer$status <- as.factor(cancer$status)
cancer$sex <- as.factor(cancer$sex)
cancer$ph.ecog <- as.factor(cancer$ph.ecog)
summary(cancer)
```
NA : inst, ph.karno, pat.karno, meal.cal, wt.loss

## Exploration of numeric data 

```{r}
# Grouping the numeric variables
num <- c("time","age","ph.karno","pat.karno","meal.cal","wt.loss")
```


```{r}
# Create repeated histograms for each categorical variable
for(var in num) {
  hist(cancer[[var]], main = paste("Histogram of", var), xlab = var)
}
```
```{r}
# Create repeated plots for each numerical variable
for(var in num) {
  plot(cancer[[var]], main = paste("Plot of", var), xlab = "Index", ylab = var)
}
```
```{r}
# Normality test using Saphiro and QQ-plot
for(var in num) {
  # Perform the Shapiro-Wilk test
shapiro_result <- shapiro.test(cancer[[var]])

# Print the result with a title
cat("Shapiro-Wilk Test Result for", var, ":\n")
print(shapiro_result)

# Create the Q-Q plot
qqnorm(cancer[[var]], main = paste("Plot of", var), xlab = "Index")
qqline(cancer[[var]], col = "red")
}
```
note : all numeric variables are not normally distributed


```{r}
#Grouping the categorical variables
cat <- cancer[,c("inst", "status","sex", "ph.ecog")]
```

```{r}
# Describe the categorical data
describe(cat)
```
```{r}
#Provide percentage of factor variable
epiDisplay::tab1(cancer$sex, graph=TRUE)

epiDisplay::tab1(cancer$status, graph=TRUE)
```
60,5% parients is male
72,4% patients dead and 27,6% censored

```{r}
#Frequency Table of Sex

# Convert 'sex' variable to factor with custom labels
cancer$sex <- factor(cancer$sex, levels = c(1, 2), labels = c("Male", "Female"))

#Convert 'status' variable to factor
cancer$status <- factor(cancer$status, levels = c(1,2), labels = c("censored","dead"))

epiDisplay::tab1(cancer$sex, 
main="Frequency Distribution of sex",
col=c("green", "yellow"), 
font.lab=2, 
font.axis=2)

epiDisplay::tab1(cancer$status, 
main="Frequency Distribution of status",
col=c("green", "yellow"), 
font.lab=2, 
font.axis=2)

```
```{r}
#Pivot table for Gender (rows) and Section (column)
library(pivottabler)

pivottabler::qpvt(cancer,"sex","status",
                  c("Mean age"="mean(age, na.rm=TRUE)",
                    "SD age"="sd(age, na.rm=TRUE)",
                    "Median age"="median(age, na.rm=TRUE)",
                    "Min age"="min(age, na.rm=TRUE)",
                    "Max age"="max(age, na.rm=TRUE)"),
                  formats=list("%.0f","%.1f"))
```
Mean of age :62(9.1) or median 63(39-82)
Mean age of female 61(8,8) and median 61(41-7)
Mean ge of male 63(9.1) and median 64(39-82)

```{r}
library(s20x)
#Summary stat of Lbs by Section
s20x::summaryStats(time~status, cancer,
                   na.rm=TRUE)
```
Mean time to event (dead) is 283 days SD 202.81 days
Mean time to censored is 363,5 days SD 221,14 days


```{r}
#Change the dataset to tibble
as_tibble(cancer)
cancer <- as_tibble(cancer)
cancer
```

```{r}
summary(cancer)
```


# Survival Curves

```{r}
# Create survival curve
s <- Surv(cancer$time, cancer$status)
```

```{r}
# Fit the survival analysis 1=censored, 2=dead
survfit(s~1)
survfit(Surv(time, status)~1, data=cancer)
sfit <- survfit(Surv(time, status)~1, data=cancer)
sfit
```

```{r}
summary(sfit)
```

```{r}
# Compare the survival time between male and female
sfit <- survfit(Surv(time, status)~sex, data=cancer)
sfit
summary(sfit)
```

```{r}
# ?summary.survfit
range(cancer$time)
seq(0, 1100, 100)
```

```{r}
summary(sfit, times=seq(0, 1000, 100))
```

Kaplan-Meier Plots

```{r}
sfit <- survfit(Surv(time, status)~sex, data=cancer)
plot(sfit)
```

```{r}
library(survminer)
ggsurvplot(sfit)
```

```{r}
ggsurvplot(sfit, conf.int=TRUE, pval=TRUE, risk.table=TRUE, 
           legend.labs=c("Male", "Female"), legend.title="Sex",  
           palette=c("dodgerblue2", "orchid2"), 
           title="Kaplan-Meier Curve for cancer Cancer Survival", 
           risk.table.height=.15)
```

```{r}
# Fit a Cox proportional hazards model
cox_fit <- coxph(Surv(time, status) ~ age + sex + ph.ecog, data=cancer)

# Summarize the Cox model
summary(cox_fit)
```
```{r}
# Visualize the Cox proportional hazards model results
ggforest(cox_fit, data = cancer)
```

